---
title: "FAIR: Data Cleaning and QC Checks"
output:
  html_document:
    toc: true
    toc_float: true
---
    
```{r setup, include=FALSE}
lib_paths <- .libPaths()
lib_paths <- c("//win.ad.jhu.edu/Users$/HOME/R/win-library/4.0", lib_paths)
.libPaths(lib_paths)
knitr::opts_chunk$set(message=FALSE, warning=FALSE, fig.align="center", fig.width=12)
```

# Load packages and read data

Load packages and utility functions.

```{r message=FALSE}
library(SummarizedExperiment)
library(matrixStats)
library(readxl)
library(tidyverse)
library(lubridate)

source("functions_utils.R")
source("functions_qc.R")
```

Read in metabolomics data.

```{r}
se_c8_fair <- read_broad_data(
    file = "../data/Hopkins FAIR Study Data/Broad/22_0825_uremic_symptoms_C8-pos.xlsx", 
    range_abund = "I10:PI5127", 
    range_row_data = "A9:H5127", 
    range_col_data = "A1:I418",
    injection = "C8"
)

se_hilic_pos_fair <- read_broad_data(
    file = "../data/Hopkins FAIR Study Data/Broad/22_0825_uremic_symptoms_HILIC-pos.xlsx", 
    range_abund = "H10:PH9654", 
    range_row_data = "A9:G9654", 
    range_col_data = "A1:I418",
    injection = "HILIC-pos"
)

se_hilic_neg_fair <- read_broad_data(
    file = "../data/Hopkins FAIR Study Data/Broad/22_0825_uremic_symptoms_HILIC-neg.xlsx", 
    range_abund = "I10:PK2071", 
    range_row_data = "A9:H2071", 
    range_col_data = "A1:I420",
    injection = "HILIC-neg"
)
```

Split `SummarizedExperiment` objects into main study samples and QC samples.

```{r}
se_c8_fair_split <- split_experiment(se_c8_fair)
se_hilic_pos_fair_split <- split_experiment(se_hilic_pos_fair)
se_hilic_neg_fair_split <- split_experiment(se_hilic_neg_fair)

se_c8_fair_main <- se_c8_fair_split$main
se_hilic_pos_fair_main <- se_hilic_pos_fair_split$main
se_hilic_neg_fair_main <- se_hilic_neg_fair_split$main
```

Read in outcome and covariate data.

```{r}
fair <- read_excel("../data/Hopkins FAIR Study Data/Original.Data from Olivia/Updated.05042022/FAIR Metabolomics_KDQOL_20220504.xlsx", range = "A1:GK374")
```





# Check sample sizes and IDs

Check sample sizes.

```{r}
dim(fair)
dim(se_c8_fair_main)
dim(se_hilic_pos_fair_main)
dim(se_hilic_neg_fair_main)
```

There is one more sample in the covariate data than in the Broad data. What is the mismatch?

```{r}
setdiff(fair$fairid, colData(se_c8_fair_main)$sample_id)
setdiff(fair$fairid, colData(se_hilic_pos_fair_main)$sample_id)
setdiff(fair$fairid, colData(se_hilic_neg_fair_main)$sample_id)
```





# Summary statistics across covariates

`hea5a`: MI
`hea5b`: peripheral vascular disease
`hea5c`: cerebral vascular disease
`hea5h`: diabetes
`hea5o`: Congestive heart failure

`lab_value12`: albumin (g/dL)
`lab_value10`: HGB (g/dL)
`lab_value15`: FERRITIN (ng/mL)
`lab_value14`: PHOS (mg/dL)
`lab_value16`: PTH (pg/mL)

```{r}
fair <- fair %>%
    dplyr::rename(
        mi = hea5a,
        pvd = hea5b,
        cvd = hea5c,
        diabetes = hea5h,
        chf = hea5o,
        alb = lab_value12,
        hgb = lab_value10,
        ferritin = lab_value15,
        phosph = lab_value14,
        ipth = lab_value16
    ) %>% 
    mutate(cv_disease = mi=="Yes" | pvd=="Yes" | cvd=="Yes" | chf=="Yes")

variables <- c("age", "sex", "race_eth", "ht_cm", "wt_kg", "bmi", "mi", "pvd", "cvd", "chf", "cv_disease", "diabetes", "alb", "hgb", "ferritin", "phosph", "ipth")
variable_types <- c("q", "c", "c", rep("q", 3), rep("c", 6), rep("q", 5))

for (i in seq_along(variables)) {
    var <- variables[i]
    cat(var, "-------------------------------------------:\n")
    if (variable_types[i]=="c")
        fair %>% count(.data[[var]]) %>% print()
    else {
        print(summary(fair[[var]]))
    }
}
```





# Cleaning symptom information

## Select and rename symptoms of interest

"During the past 4 weeks, to what extent were you bothered by..."

Main symptoms

- Fatigue: Being washed out or drained?
    - `qol14i`: 14.I. Washed out or drained?
- Pruritus: Itchy skin?
    - `qol14d`: 14.D. Itchy skin?
- Anorexia: Lack of appetite?
    - `qol14h`: 14.H. Lack of appetite?

Secondary symptoms

- Nausea/Vomiting: Nausea or upset stomach?
    - `qol14k`: 14.K. Nausea or upset stomach?
- Excessive daytime sleep: Sleepiness during the day?
    - `qol18c`: 18.C. Have trouble staying awake during the day?
- Difficulty concentrating: Difficulty doing activities involving concentrating and thinking?
    - `qol13d`: 13.D. Did you have difficulty concentrating or thinking?
- Pain: Bodily pain?
    - `qol7`: 7. How much bodily pain have you had during the past 4 weeks? 

Extra variables related to fatigue

- Climbing one flight of stairs
    - `qol3e`: 3.E. Climbing one flight of stairs.
- Climbing several flights of stairs
    - `qol3d`: 3.D. Climbing several flights of stairs. 
- Moderate activity
    - `qol3b`: 3.B. Moderate activities, such as moving a table, pushing a vacuum cleaner, bowl
- Vigorous activity
    - `qol3a`: 3.A. Vigorous activities, such as running, lifting heavy objects, participating 
- Walking one block
    - `qol3i`: 3.I. Walking one block.
- Walking several blocks
    - `qol3h`: 3.H. Walking several blocks.
- Walking a mile or more
    - `qol3g`: 3.G. Walking more than a mile.

Rename (abbreviate) variables for symptoms of interest.

```{r}
fair <- fair %>%
    dplyr::rename(
        fatigue = qol14i,
        pruritus = qol14d,
        anorexia = qol14h,
        nausea_vomiting = qol14k,
        daytime_sleepiness = qol18c,
        difficulty_concentrating = qol13d,
        bodily_pain = qol7,
        climb_one_flight = qol3e,
        climb_several_flights = qol3d,
        moderate_activity = qol3b,
        vigorous_activity = qol3a,
        walk_one_block = qol3i,
        walk_several_blocks = qol3h,
        walk_mile_more = qol3g
    )
```

What are the possible responses for the selected symptoms?

```{r}
fair %>% dplyr::count(fatigue)
fair %>% dplyr::count(pruritus)
fair %>% dplyr::count(anorexia)
fair %>% dplyr::count(nausea_vomiting)
fair %>% dplyr::count(daytime_sleepiness)
fair %>% dplyr::count(difficulty_concentrating)
fair %>% dplyr::count(bodily_pain)

fair %>% dplyr::count(climb_one_flight)
fair %>% dplyr::count(climb_several_flights)
fair %>% dplyr::count(moderate_activity)
fair %>% dplyr::count(vigorous_activity)
fair %>% dplyr::count(walk_one_block)
fair %>% dplyr::count(walk_several_blocks)
fair %>% dplyr::count(walk_mile_more)
```

## Recode categories for symptoms

```{r}
recode_categories <- function(x, type = c("type1", "type2", "type3", "type4")) {
    if (type=="type1") {
        case_when(
            x=="1. Not at all bothered" ~ "grade1",
            x=="2. Somewhat bothered" ~ "grade2",
            x=="3. Moderately bothered" ~ "grade3",
            x=="4. Very much bothered" ~ "grade4",
            x=="5. Extremely bothered" ~ "grade5"
        )
    } else if (type=="type2") {
        case_when(
            x=="1. None of the time" ~ "grade1",
            x=="2. A little of the time" ~ "grade2",
            x=="3. Some of the time" ~ "grade3",
            x=="4. A good bit of the time" ~ "grade4",
            x=="5. Most of the time" ~ "grade5",
            x=="6. All of the time" ~ "grade6"
        )
    } else if (type=="type3") {
        case_when(
            x=="1. None" ~ "grade1",
            x=="2. Very mild" ~ "grade2",
            x=="3. Mild" ~ "grade3",
            x=="4. Moderate" ~ "grade4",
            x=="5. Severe" ~ "grade5",
            x=="6. Very Severe" ~ "grade6"
        )
    } else if (type=="type4") {
        case_when(
            x=="3. No, not limited at all" ~ "grade1",
            x=="2. Yes, limited a little" ~ "grade2",
            x=="1. Yes, limited a lot" ~ "grade3"
        )
    }
}

fair_new <- fair %>%
    mutate(across(c(fatigue, pruritus, anorexia, nausea_vomiting), recode_categories, type = "type1")) %>%
    mutate(across(c(daytime_sleepiness, difficulty_concentrating), recode_categories, type = "type2")) %>%
    mutate(bodily_pain = recode_categories(bodily_pain, type = "type3")) %>%
    mutate(across(c(climb_one_flight, climb_several_flights, moderate_activity, vigorous_activity, walk_one_block, walk_several_blocks, walk_mile_more), recode_categories, type = "type4"))
```





# Merge covariate data into SummarizedExperiments

```{r}
merge_se_covars <- function(se, df_covariates) {
    cd <- colData(se) %>% as.data.frame()
    cd <- cd %>%
        left_join(df_covariates, by = c("sample_id" = "fairid"))
    colData(se) <- as(cd, "DataFrame")
    se
}

se_c8_fair_main <- merge_se_covars(se_c8_fair_main, fair_new)
se_hilic_pos_fair_main <- merge_se_covars(se_hilic_pos_fair_main, fair_new)
se_hilic_neg_fair_main <- merge_se_covars(se_hilic_neg_fair_main, fair_new)
```





# QC: Sample drift and batch effects

## Sample drift for QC samples

```{r}
plot_qc_drift(se_c8_fair_split$main, sample_descrip = "C8 - Main", include_subtitle_corr = TRUE)
plot_qc_drift(se_c8_fair_split$qc_a, sample_descrip = "C8 - PREFA", include_subtitle_corr = TRUE)
plot_qc_drift(se_c8_fair_split$qc_b, sample_descrip = "C8 - PREFB", include_subtitle_corr = TRUE)
plot_qc_drift(se_c8_fair_split$lucid_pool, sample_descrip = "C8 - LUCID pool", include_subtitle_corr = TRUE)

plot_qc_drift(se_hilic_pos_fair_split$main, sample_descrip = "HILIC-pos - Main", include_subtitle_corr = TRUE)
plot_qc_drift(se_hilic_pos_fair_split$qc_a, sample_descrip = "HILIC-pos - PREFA", include_subtitle_corr = TRUE)
plot_qc_drift(se_hilic_pos_fair_split$qc_b, sample_descrip = "HILIC-pos - PREFB", include_subtitle_corr = TRUE)
plot_qc_drift(se_hilic_pos_fair_split$lucid_pool, sample_descrip = "HILIC-pos - LUCID pool", include_subtitle_corr = TRUE)

plot_qc_drift(se_hilic_neg_fair_split$main, sample_descrip = "HILIC-neg - Main", include_subtitle_corr = TRUE)
plot_qc_drift(se_hilic_neg_fair_split$qc_a, sample_descrip = "HILIC-neg - PREFA", include_subtitle_corr = TRUE)
plot_qc_drift(se_hilic_neg_fair_split$qc_b, sample_descrip = "HILIC-neg - PREFB", include_subtitle_corr = TRUE)
plot_qc_drift(se_hilic_neg_fair_split$lucid_pool, sample_descrip = "HILIC-neg - LUCID pool", include_subtitle_corr = TRUE)
```


## Coefficient of variation distributions

```{r}
plot_cv_dists(se_c8_fair_split, "C8", xlim = c(0,1))
plot_cv_dists(se_hilic_pos_fair_split, "HILIC (pos)", xlim = c(0,10))
plot_cv_dists(se_hilic_neg_fair_split, "HILIC (neg)", xlim = c(0,10))
```


## PCA

### C8

```{r}
plot_pc_info(se_c8_fair_split$main, sample_descrip = "C8: Main")
plot_pc_info(se_c8_fair_split$qc_a, sample_descrip = "C8: PREFA")
plot_pc_info(se_c8_fair_split$qc_b, sample_descrip = "C8: PREFB")
```


### HILIC-pos

```{r}
plot_pc_info(se_hilic_pos_fair_split$main, sample_descrip = "HILIC (pos): Main")
plot_pc_info(se_hilic_pos_fair_split$qc_a, sample_descrip = "HILIC (pos): PREFA")
plot_pc_info(se_hilic_pos_fair_split$qc_b, sample_descrip = "HILIC (pos): PREFB")
```


### HILIC-neg

```{r}
plot_pc_info(se_hilic_neg_fair_split$main, sample_descrip = "HILIC (neg): Main")
plot_pc_info(se_hilic_neg_fair_split$qc_a, sample_descrip = "HILIC (neg): PREFA")
plot_pc_info(se_hilic_neg_fair_split$qc_b, sample_descrip = "HILIC (neg): PREFB")
```





# Metabolite redundancy across injections

Read in Eugene's annotations.

```{r}
metab_redundancy <- read_excel("../data/Broad data--redundant metabolites_02032022.xlsx", sheet = "Sheet1", range = "A1:I665", col_names = TRUE)
colnames(metab_redundancy) <- fix_names(colnames(metab_redundancy))
metab_redundancy <- metab_redundancy %>% select(-`...6`)
```

What metabolites are measured in multiple injections?

```{r}
metabs_c8 <- setdiff(rowData(se_c8_fair_main)$metabolite, NA)
metabs_hilic_pos <- setdiff(rowData(se_hilic_pos_fair_main)$metabolite, NA)
metabs_hilic_neg <- setdiff(rowData(se_hilic_neg_fair_main)$metabolite, NA)
df_metabs_by_method <- tibble(
    all_metabs = unique(c(metabs_c8, metabs_hilic_pos, metabs_hilic_neg)),
    c8 = all_metabs %in% metabs_c8,
    hil_pos = all_metabs %in% metabs_hilic_pos,
    hil_neg = all_metabs %in% metabs_hilic_neg
)

df_metabs_by_method[rowSums(df_metabs_by_method[,2:4]) > 1,]
```

Remove redundant metabolites across injections.

- **Note:** the `metabolite` column contains new nomenclature, so the metabolite names in this column don't match Eugene's annotation from the LUCID files. The C8 and HILIC-neg injections have a `former_nomenclature` column that 

```{r}
se_c8_fair_main <- remove_metabolites(se_c8_fair_main, metab_redundancy, injection = "C8-pos", row_data_col = "former_nomenclature")
se_hilic_pos_fair_main <- remove_metabolites(se_hilic_pos_fair_main, metab_redundancy, injection = "HIL-pos", row_data_col = "metabolite")
se_hilic_neg_fair_main <- remove_metabolites(se_hilic_neg_fair_main, metab_redundancy, injection = "HIL-neg", row_data_col = "former_nomenclature")
```


# Save data before filtering metabolites

```{r}
save(se_c8_fair_main, se_hilic_pos_fair_main, se_hilic_neg_fair_main, file = "../data/summ_exps_fair_clean_pre_filter.rda")
```




# Filtering metabolites based on missingness and zero variance

- Remove metabolites with >= 95% missingness
- Remove metabolites where the variance in abundance among non-missing values is zero

```{r}
se_c8_fair_main <- filter_metabs(se_c8_fair_main)
se_hilic_pos_fair_main <- filter_metabs(se_hilic_pos_fair_main)
se_hilic_neg_fair_main <- filter_metabs(se_hilic_neg_fair_main)
```





# Save clean data

```{r}
save(se_c8_fair_main, se_hilic_pos_fair_main, se_hilic_neg_fair_main, file = "../data/summ_exps_fair_clean.rda")
```




