---
title: "LUCID: Symptom analysis"
output:
  html_document:
    toc: true
    toc_float: true
---
    
```{r setup, include=FALSE}
lib_paths <- .libPaths()
lib_paths <- c("//win.ad.jhu.edu/Users$/HOME/R/win-library/4.0", lib_paths)
.libPaths(lib_paths)
knitr::opts_chunk$set(message=FALSE, warning=FALSE, fig.align="center")
```

# Notes

- The `cv_disease` and `diabetes` variables represent history of **EVER** having a positive diagnosis.
    - CV disease:
        - US: CAD, cardiac arrest, cardiac arrhythmia, MI, stroke
        - Canada: Same as US but not stroke
        
Order in which Rmd files must be run:

1. `lucid_cleaning_qc.Rmd` / `fair_cleaning_qc.Rmd`
2. `imputation_ruv.Rmd` / `fair_imputation_ruv.Rmd`
3. `lucid_main_analysis.Rmd`





# Load packages

Load packages and utility functions.

```{r message=FALSE}
library(SummarizedExperiment)
library(matrixStats)
library(readxl)
library(tidyverse)
library(ggthemes)
library(lubridate)
library(limma)
library(splines)
library(caret)
library(glmnet)
library(randomForest)

source("functions_utils.R")
source("functions_limma_analysis.R")
source("functions_ml.R")
```

# Metabolite counting

Load data where original number of metabolites can be counted

```{r}
load("../data/summ_exps_clean_pre_filter.rda")
load("../data/summ_exps_fair_clean_pre_filter.rda")
```

Remove internal standards

```{r}
# LUCID
se_c8_main <- remove_internal_standards(se_c8_main)
se_hilic_pos_main <- remove_internal_standards(se_hilic_pos_main)
se_hilic_neg_main <- remove_internal_standards(se_hilic_neg_main)

# FAIR
se_c8_fair_main <- remove_internal_standards(se_c8_fair_main)
se_hilic_pos_fair_main <- remove_internal_standards(se_hilic_pos_fair_main)
se_hilic_neg_fair_main <- remove_internal_standards(se_hilic_neg_fair_main)
```

```{r}
rd_lucid <- bind_rows(
    rowData(se_c8_main) %>% as.data.frame() %>% mutate(hmdb_id_certainty = as.character(hmdb_id_certainty)),
    rowData(se_hilic_pos_main) %>% as.data.frame() %>% mutate(hmdb_id_certainty = as.character(hmdb_id_certainty)),
    rowData(se_hilic_neg_main) %>% as.data.frame() %>% mutate(hmdb_id_certainty = as.character(hmdb_id_certainty))
)
rd_fair <- bind_rows(
    rowData(se_c8_fair_main) %>% as.data.frame(),
    rowData(se_hilic_pos_fair_main) %>% as.data.frame(),
    rowData(se_hilic_neg_fair_main) %>% as.data.frame()
)
```

Get unique known metabolites across LUCID and FAIR.

```{r}
setdiff(c(rd_lucid$hmdb_id, rd_fair$hmdb_id), NA) %>% length()
```

Get unknowns.

```{r}
unknowns_lucid <- rd_lucid %>%
    filter(is.na(hmdb_id)) %>% 
    select(method, mz, rt)
unknowns_fair <- rd_fair %>%
    filter(is.na(hmdb_id)) %>% 
    select(method, mz, rt)
unknowns <- bind_rows(unknowns_lucid, unknowns_fair)
dim(unknowns)
dim(unique(unknowns))

round_any <- function(x, accuracy, f = round) {
    f(x/accuracy)*accuracy
}

unknowns <- unknowns %>% 
    mutate(mz = round_any(mz, .05), rt = round_any(rt, 0.25))
dim(unique(unknowns))
```

Remove this data to avoid conflicts with loading in clean data.

```{r}
rm(se_c8_main, se_hilic_pos_main, se_hilic_neg_main)
rm(se_c8_fair_main, se_hilic_pos_fair_main, se_hilic_neg_fair_main)
```




# Load clean data for analyses

```{r}
load("../data/summ_exps_imputed_ruv.rda")
load("../data/summ_exps_fair_imputed_ruv.rda")
```





# Data cleaning

## Remove internal standards

LUCID

```{r}
se_c8_main <- remove_internal_standards(se_c8_main)
se_hilic_pos_main <- remove_internal_standards(se_hilic_pos_main)
se_hilic_neg_main <- remove_internal_standards(se_hilic_neg_main)
```

FAIR

```{r}
se_c8_fair_main <- remove_internal_standards(se_c8_fair_main)
se_hilic_pos_fair_main <- remove_internal_standards(se_hilic_pos_fair_main)
se_hilic_neg_fair_main <- remove_internal_standards(se_hilic_neg_fair_main)
```

## Update `site` variable

```{r}
se_c8_main <- update_site_variable(se_c8_main)
se_hilic_pos_main <- update_site_variable(se_hilic_pos_main)
se_hilic_neg_main <- update_site_variable(se_hilic_neg_main)
```

## Add `former_nomenclature` column to HILIC+ in FAIR

```{r}
rowData(se_hilic_pos_fair_main)$former_nomenclature <- NA
```

## Check overlaps in HMDB IDs across LUCID and FAIR

```{r}
check_lucid_fair_overlap <- function(se_lucid, se_fair) {
    ids_lucid <- setdiff(rowData(se_lucid)$hmdb_id, NA)
    ids_fair <- setdiff(rowData(se_fair)$hmdb_id, NA)
    ov <- intersect(ids_lucid, ids_fair)
    cat("Unique HMDB IDs in LUCID:", length(ids_lucid), "\n")
    cat("Unique HMDB IDs in FAIR:", length(ids_fair), "\n")
    cat("# in common:", length(ov), "\n")
}
check_lucid_fair_overlap(se_c8_main, se_c8_fair_main)
check_lucid_fair_overlap(se_hilic_pos_main, se_hilic_pos_fair_main)
check_lucid_fair_overlap(se_hilic_neg_main, se_hilic_neg_fair_main)
```

## Relabel QI04755 in FAIR

```{r}
rowData(se_c8_fair_main) %>%
    as.data.frame() %>% 
    filter(str_detect(compound_id, "04755"))

rowData(se_hilic_pos_fair_main) %>%
    as.data.frame() %>% 
    filter(str_detect(compound_id, "04755"))

rowData(se_hilic_neg_fair_main) %>%
    as.data.frame() %>% 
    filter(str_detect(compound_id, "04755"))

rowData(se_hilic_neg_main) %>%
    as.data.frame() %>% 
    filter(str_detect(metabolite, "2-hydroxy-3-methylpentanoate"))
```

```{r}
index_lucid <- which(str_detect(rowData(se_hilic_neg_main)$metabolite, "2-hydroxy-3-methylpentanoate"))
index_fair <- which(str_detect(rowData(se_hilic_neg_fair_main)$compound_id, "04755"))

rowData(se_hilic_neg_fair_main)$hmdb_id[index_fair] <- rowData(se_hilic_neg_main)$hmdb_id[index_lucid]
rowData(se_hilic_neg_fair_main)$metabolite[index_fair] <- rowData(se_hilic_neg_main)$metabolite[index_lucid]

rowData(se_hilic_neg_fair_main) %>%
    as.data.frame() %>% 
    filter(str_detect(compound_id, "04755"))
```




# Summary statistics across covariates

LUCID

```{r}
cd <- colData(se_c8_main) %>%
    as.data.frame() %>%
    filter(event_name=="Baseline")
cd_year1 <- colData(se_c8_main) %>%
    as.data.frame() %>%
    filter(event_name=="Year 1")

# Subset to analytic sample
# mod_form <- ~symptom+ns(age,df=3)+gender+ns(height,df=3)+ns(dry_weight,df=3)+cv_disease+diabetes+country

bool_cc_baseline <- complete.cases(cd[,c("age", "gender", "height", "dry_weight", "cv_disease", "diabetes", "country")])
table(bool_cc_baseline)
cd_baseline_analytic <- cd[bool_cc_baseline,]

bool_cc_year1 <- complete.cases(cd_year1[,c("age", "gender", "height", "dry_weight", "cv_disease", "diabetes", "country")])
bool_any_symptom_year1 <- rowSums(is.na(cd_year1[,c("fatigue", "pruritus", "anorexia", "nausea_vomiting", "daytime_sleepiness", "difficulty_concentrating", "bodily_pain")])) < 7
table(bool_cc_year1)
cd_year1_analytic <- cd_year1[bool_cc_year1,]

show_summary_stats <- function(coldat) {
    variables <- c("age", "gender", "race", "height", "dry_weight", "bmi", "cad", "cardiac_arrest", "cardiac_arrhythmias", "mi", "stroke", "cv_disease", "diabetes", "alb", "hgb", "ferritin", "phosph", "ipth")
    variable_types <- c("q", "c", "c", rep("q", 3), rep("c", 7), rep("q", 5))
    
    for (i in seq_along(variables)) {
        var <- variables[i]
        cat(var, "-------------------------------------------:\n")
        if (variable_types[i]=="c") {
            coldat %>% count(.data[[var]]) %>% mutate(frac = n/sum(n)) %>% print()
            coldat %>% count(country, .data[[var]]) %>% print()
        } else {
            print(summary(coldat[[var]]))
        }
    }
}

show_summary_stats(cd_baseline_analytic)
show_summary_stats(cd_year1_analytic)

symptoms <- c("fatigue", "pruritus", "anorexia", "nausea_vomiting", "daytime_sleepiness", "difficulty_concentrating", "bodily_pain")

for (event in c("Baseline", "Year 1")) {
    if (event=="Baseline") {
        cd <- cd_baseline_analytic
    } else if (event=="Year 1") {
        cd <- cd_year1_analytic
    }
    cat("*****", event, "*****\n")
    for (i in seq_along(symptoms)) {
        symp <- symptoms[i]
        n_this_symp <- sum(!is.na(cd[[symp]]))
        n_na_this_symp <- sum(is.na(cd[[symp]]))
        n_na_this_symp_us <- sum(is.na(cd[[symp]]) & cd$country=="US")
        n_na_this_symp_can <- sum(is.na(cd[[symp]]) & cd$country=="Canada")
        num_symptomatic <- sum(cd[[symp]]!="grade1", na.rm = TRUE)
        cat(symp, "-------------------------------------------:\n")
        cat("# with missing responses:", n_na_this_symp, "(US:", n_na_this_symp_us, ", Canada:", n_na_this_symp_can, ")\n")
        cat("Total non-missing responses:", n_this_symp, "\n")
        cat("# symptomatic:", num_symptomatic, "\n")
        cat("Frac symptomatic:", num_symptomatic/n_this_symp, "\n")
    }
    
    cat("# missing all symptoms:\n")
    tab1 <- table(rowSums(is.na(cd[,symptoms]))==7, useNA = "ifany")
    print(tab1)
    print(prop.table(tab1))
    tab1_by_country <- table(rowSums(is.na(cd[,symptoms]))==7, cd$country, useNA = "ifany")
    print(tab1_by_country)
    cat("# symptomatic for any symptom:\n")
    tab2 <- table(rowSums(cd[,symptoms]!="grade1", na.rm = TRUE) > 0, useNA = "ifany")
    print(tab2)
    print(prop.table(tab2))
}
```

FAIR

```{r}
cd_fair <- colData(se_c8_fair_main) %>% as.data.frame()

# Subset to analytic sample
# mod_form <- ~symptom+ns(age,df=3)+sex+ns(ht_cm,df=3)+ns(wt_kg,df=3)+cv_disease+diabetes

bool_cc_fair <- complete.cases(cd_fair[,c("age", "sex", "ht_cm", "wt_kg", "cv_disease", "diabetes")])
cd_fair_analytic <- cd_fair[bool_cc_fair,]


variables <- c("age", "sex", "race_eth", "ht_cm", "wt_kg", "bmi", "mi", "pvd" , "cvd", "chf", "cv_disease", "diabetes", "alb", "hgb", "ferritin", "phosph", "ipth")
variable_types <- c("q", "c", "c", rep("q", 3), rep("c", 6), rep("q", 5))

for (i in seq_along(variables)) {
    var <- variables[i]
    cat(var, "-------------------------------------------:\n")
    if (variable_types[i]=="c") {
        cd_fair_analytic %>% count(.data[[var]]) %>% mutate(frac = n/sum(n)) %>% print()
    } else {
        print(summary(cd_fair_analytic[[var]]))
    }
}

for (i in seq_along(symptoms)) {
    symp <- symptoms[i]
    n_this_symp <- sum(!is.na(cd_fair_analytic[[symp]]))
    num_symptomatic <- sum(cd_fair_analytic[[symp]]!="grade1", na.rm = TRUE)
    cat(symp, "-------------------------------------------:\n")
    cat("Total non-missing responses:", n_this_symp, "\n")
    cat("# symptomatic:", num_symptomatic, "\n")
    cat("Frac symptomatic:", num_symptomatic/n_this_symp, "\n")
}

cat("# missing all symptoms:\n")
tab1 <- table(rowSums(is.na(cd_fair_analytic[,symptoms]))==7, useNA = "ifany")
print(tab1)
print(prop.table(tab1))
cat("# symptomatic for any symptom:\n")
tab2 <- table(rowSums(cd_fair_analytic[,symptoms]!="grade1", na.rm = TRUE) > 0, useNA = "ifany")
print(tab2)
print(prop.table(tab2))
```





# Check missing data and sample sizes

```{r}
se_list <- list(
    c8 = se_c8_main,
    hilic_pos = se_hilic_pos_main,
    hilic_neg = se_hilic_neg_main
)
se_all <- merge_ses_across_injections(se_list)
rd <- rowData(se_all)
sum(!is.na(rd$metabolite))
cd <- colData(se_all)
```

```{r}
cols <- c("proton_id", "studynum", "event_name", "age", "gender", "height", "dry_weight", "cv_disease", "diabetes", "country")
cd_subs <- cd[,cols] %>%
    as.data.frame() %>% 
    mutate(id = if_else(country=="Canada", studynum, proton_id)) %>% 
    select(-proton_id, -studynum)
cd_baseline <- cd_subs %>% filter(event_name=="Baseline") %>% select(-event_name)
cd_year1 <- cd_subs %>% filter(event_name=="Year 1") %>% select(-event_name)
colnames(cd_baseline) <- str_c(colnames(cd_baseline), "_b")
cd_baseline <- cd_baseline %>% rename(id = id_b)
colnames(cd_year1) <- str_c(colnames(cd_year1), "_y1")
cd_year1 <- cd_year1 %>% rename(id = id_y1)

sum(complete.cases(cd_baseline))
sum(complete.cases(cd_year1))

summary(cd_baseline)
summary(cd_year1)

cd_year1_baseline <- left_join(cd_year1, cd_baseline, by = "id")
```





# Total number of metabolites across LUCID and FAIR

```{r}
se_list <- list(
    c8 = se_c8_main,
    hilic_pos = se_hilic_pos_main,
    hilic_neg = se_hilic_neg_main
)

se_list_fair <- list(
    c8 = se_c8_fair_main,
    hilic_pos = se_hilic_pos_fair_main,
    hilic_neg = se_hilic_neg_fair_main
)

se_all_lucid <- merge_ses_across_injections(se_list)
se_all_fair <- merge_ses_across_injections(se_list_fair)

rd_lucid <- rowData(se_all_lucid) %>% as.data.frame()
rd_fair <- rowData(se_all_fair) %>% as.data.frame()

setdiff(c(rd_lucid$hmdb_id, rd_fair$hmdb_id), NA) %>% length()
```

```{r}
unknowns_lucid <- rd_lucid %>%
    filter(is.na(hmdb_id)) %>% 
    select(method, mz, rt)
unknowns_fair <- rd_fair %>%
    filter(is.na(hmdb_id)) %>% 
    select(method, mz, rt)
unknowns <- bind_rows(unknowns_lucid, unknowns_fair)
dim(unknowns)
dim(unique(unknowns))

unknowns <- unknowns %>% 
    mutate(mz = round_any(mz, .05), rt = round_any(rt, 0.25))
dim(unique(unknowns))
```





# Fit ordinary regression models

## LUCID

Set up sensitivity analysis specifications.

```{r}
df_analysis_versions <- tidyr::crossing(
    metab_type = c("orig", "qrilc_no_ruv", "qrilc_ruv"),
    imputed_symptoms = c(FALSE, TRUE)
)
df_analysis_versions <- df_analysis_versions %>%
    mutate(
        version = paste0("v", seq_len(nrow(df_analysis_versions)))
    )
```

Run analyses across sensitivity analysis specifications.

```{r}
linreg_results <- vector("list", nrow(df_analysis_versions))

mod_form <- symptom ~ metab+ns(age,df=3)+gender+ns(height,df=3)+ns(dry_weight,df=3)+cv_disease+diabetes+country

for (i in seq_len(nrow(df_analysis_versions))) {
    linreg_results[[i]] <- run_all_linreg_analyses(
        se_list = se_list,
        mod_form = mod_form,
        cohort = "LUCID",
        metab_type = df_analysis_versions$metab_type[i],
        use_imputed_symptoms = df_analysis_versions$imputed_symptoms[i]
    )
}

linreg_results_lucid_df_list <- lapply(linreg_results, results_list_to_df)
names(linreg_results_lucid_df_list) <- df_analysis_versions$version
df_linreg_results_lucid <- bind_rows(linreg_results_lucid_df_list, .id = "version") %>%
    mutate(cohort = "LUCID") %>%
    left_join(df_analysis_versions, by = "version")
```

## FAIR

Set up axes of sensitivity analysis for FAIR.

```{r}
df_analysis_versions <- tidyr::crossing(
    metab_type = c("orig", "qrilc_no_ruv", "qrilc_ruv")
)
df_analysis_versions <- df_analysis_versions %>%
    mutate(
        version = paste0("v", seq_len(nrow(df_analysis_versions)))
    )
```

Run analyses across sensitivity analysis specifications.

```{r}
linreg_results_fair <- vector("list", nrow(df_analysis_versions))

mod_form <- symptom ~ metab+ns(age,df=3)+sex+ns(ht_cm,df=3)+ns(wt_kg,df=3)+cv_disease+diabetes

for (i in seq_len(nrow(df_analysis_versions))) {
    linreg_results_fair[[i]] <- run_all_linreg_analyses(
        se_list = se_list_fair,
        mod_form = mod_form,
        cohort = "FAIR",
        metab_type = df_analysis_versions$metab_type[i],
        use_imputed_symptoms = FALSE
    )
}

linreg_results_fair_df_list <- lapply(linreg_results_fair, results_list_to_df)
names(linreg_results_fair_df_list) <- df_analysis_versions$version
df_linreg_results_fair <- bind_rows(linreg_results_fair_df_list, .id = "version") %>%
    mutate(cohort = "FAIR") %>%
    left_join(df_analysis_versions, by = "version")

## Change "HILIC" to "HIL" in `method` variable
df_linreg_results_fair <- df_linreg_results_fair %>%
    mutate(method = str_replace(method, "HILIC", "HIL"))
```












# Fit limma models

## LUCID

Set up axes of sensitivity analyses.

- Removed "race" and "ethnicity" for now due to problems estimating those coefficients (just some of the race coefficients actually, but removing ethnicity due to removing race)

**REVISIT:** Problems when adjusting for site. `Coefficients not estimable: siteSeattle`

"age", "gender", "race", "height", "dry_weight", "bmi", "cad", "cardiac_arrest", "cardiac_arrhythmias", "mi", "stroke", "diabetes", "alb", "hgb", "ferritin", "phosph", "ipth"

```{r}
# mod_forms <- c(
#     ~symptom+ns(age,df=3)+gender+ns(height,df=3)+ns(dry_weight,df=3)+cv_disease+diabetes,
#     ~symptom+ns(age,df=3)+gender+ns(height,df=3)+ns(dry_weight,df=3)+cv_disease+diabetes+ns(alb_imp,df=3)+ns(hgb_imp,df=3)+ns(ferritin_imp,df=3)+ns(phosph_imp,df=3)+ns(ipth_imp,df=3)
# )

df_analysis_versions <- tidyr::crossing(
    # mod_form = mod_forms,
    metab_type = c("orig", "qrilc_no_ruv", "qrilc_ruv"),
    imputed_symptoms = c(FALSE, TRUE)
)
df_analysis_versions <- df_analysis_versions %>%
    mutate(
        version = paste0("v", seq_len(nrow(df_analysis_versions)))
    )
```

Run limma analyses across sensitivity analysis specifications.

```{r}
se_list <- list(
    c8 = se_c8_main,
    hilic_pos = se_hilic_pos_main,
    hilic_neg = se_hilic_neg_main
)

limma_results <- vector("list", nrow(df_analysis_versions))

mod_form <- ~symptom+ns(age,df=3)+gender+ns(height,df=3)+ns(dry_weight,df=3)+cv_disease+diabetes+country

for (i in seq_len(nrow(df_analysis_versions))) {
    limma_results[[i]] <- run_all_analyses(
        se_list = se_list,
        mod_form = mod_form,
        cohort = "LUCID",
        coarse_symptom_levels = 3,
        metab_type = df_analysis_versions$metab_type[i],
        use_imputed_symptoms = df_analysis_versions$imputed_symptoms[i]
    )
}

limma_results_lucid_df_list <- lapply(limma_results, results_list_to_df)
names(limma_results_lucid_df_list) <- df_analysis_versions$version
df_limma_results_lucid <- bind_rows(limma_results_lucid_df_list, .id = "version") %>%
    mutate(cohort = "LUCID") %>%
    left_join(df_analysis_versions, by = "version")
```

## FAIR

Set up axes of sensitivity analysis for FAIR.

```{r}
df_analysis_versions <- tidyr::crossing(
    metab_type = c("orig", "qrilc_no_ruv", "qrilc_ruv")
)
df_analysis_versions <- df_analysis_versions %>%
    mutate(
        version = paste0("v", seq_len(nrow(df_analysis_versions)))
    )
```

Run limma analyses across sensitivity analysis specifications.

```{r}
se_list_fair <- list(
    c8 = se_c8_fair_main,
    hilic_pos = se_hilic_pos_fair_main,
    hilic_neg = se_hilic_neg_fair_main
)

limma_results_fair <- vector("list", nrow(df_analysis_versions))

mod_form <- ~symptom+ns(age,df=3)+sex+ns(ht_cm,df=3)+ns(wt_kg,df=3)+cv_disease+diabetes

for (i in seq_len(nrow(df_analysis_versions))) {
    limma_results_fair[[i]] <- run_all_analyses(
        se_list = se_list_fair,
        mod_form = mod_form,
        cohort = "FAIR",
        coarse_symptom_levels = 3,
        metab_type = df_analysis_versions$metab_type[i],
        use_imputed_symptoms = FALSE
    )
}

limma_results_fair_df_list <- lapply(limma_results_fair, results_list_to_df)
names(limma_results_fair_df_list) <- df_analysis_versions$version
df_limma_results_fair <- bind_rows(limma_results_fair_df_list, .id = "version") %>%
    mutate(cohort = "FAIR") %>%
    left_join(df_analysis_versions, by = "version")

## Change "HILIC" to "HIL" in `method` variable
df_limma_results_fair <- df_limma_results_fair %>%
    mutate(method = str_replace(method, "HILIC", "HIL"))
```





# Fit ML models

## LUCID

```{r}
adj_vars <- c("age", "gender", "height", "dry_weight", "cv_disease", "diabetes", "country") # "race", "alb", "hgb", "ferritin", "phosph", "ipth"

df_analysis_versions <- tidyr::crossing(
    metab_type = c("qrilc_no_ruv", "qrilc_ruv"),
    imputed_symptoms = c(FALSE, TRUE)
)
df_analysis_versions <- df_analysis_versions %>%
    mutate(
        version = paste0("v", seq_len(nrow(df_analysis_versions)))
    )

set.seed(248)
ml_results_lucid <- lapply(seq_len(nrow(df_analysis_versions)), function(i) {
    run_all_ml_analyses(
        se_list = se_list,
        adjustment_vars = adj_vars,
        cohort = "LUCID",
        only_known_metabs = TRUE,
        coarse_symptom_levels = 3,
        metab_type = df_analysis_versions$metab_type[i],
        use_imputed_symptoms = df_analysis_versions$imputed_symptoms[i]
    )
})

ml_results_lucid_df_list  <- lapply(ml_results_lucid, ml_results_to_df)
names(ml_results_lucid_df_list) <- df_analysis_versions$version
df_ml_results_lucid <- bind_rows(ml_results_lucid_df_list, .id = "version") %>%
    mutate(cohort = "LUCID") %>%
    left_join(df_analysis_versions, by = "version") %>%
    filter(!(metab_id %in% adj_vars))
```

Plots of test accuracy vs. tuning parameters

```{r}
par(mfrow = c(1,2))
for (i in seq_along(ml_results_lucid)) {
    list_this_sens_version <- ml_results_lucid[[i]]
    for (event in names(list_this_sens_version)) {
        list_this_event <- list_this_sens_version[[event]]
        for (symp in names(list_this_event)) {
            list_this_symptom <- list_this_event[[symp]]
            plot(list_this_symptom[[1]], main = paste0("Version ", i, ". Event: ", event, ". Symptom: ", symp))
            plot(list_this_symptom[[2]], main = paste0("Version ", i, ". Event: ", event, ". Symptom: ", symp))
        }
    }
}
```


## FAIR

```{r}
adj_vars <- c("age", "sex", "ht_cm", "wt_kg", "cv_disease", "diabetes")
metab_types <- c("qrilc_no_ruv", "qrilc_ruv")

set.seed(282)
ml_results_fair <- lapply(metab_types, function(metab_type) {
    run_all_ml_analyses(
        se_list = se_list_fair,
        adjustment_vars = adj_vars,
        cohort = "FAIR",
        only_known_metabs = TRUE,
        coarse_symptom_levels = 3,
        metab_type = metab_type,
        use_imputed_symptoms = FALSE
    )
})

ml_results_fair_df_list  <- lapply(ml_results_fair, ml_results_to_df)
names(ml_results_fair_df_list) <- metab_types
df_ml_results_fair <- bind_rows(ml_results_fair_df_list, .id = "metab_type") %>%
    mutate(cohort = "FAIR") %>%
    filter(!(metab_id %in% adj_vars))
```

Save results.

```{r}
save(df_linreg_results_lucid, df_linreg_results_fair, df_limma_results_lucid, df_limma_results_fair, df_ml_results_lucid, df_ml_results_fair, file = "../data/symptoms_results_df.rda")
```





# Look at QI04755

```{r}
df_linreg_results_fair %>% filter(compound_id=="QI04755")
df_limma_results_fair %>% filter(compound_id=="QI04755")
```

```{r}
df_linreg_results_fair %>% filter(compound_id=="QI04755") %>% arrange(p.value)
df_limma_results_fair %>% filter(compound_id=="QI04755") %>% arrange(P.Value)
```

# Plots and tables

For "interesting" metabolites, want to know:

- `limma` results: Significant association?
    - At baseline? At Year 1? In FAIR?
- ML results: At least medium importance in LASSO? In RF?
    - At baseline? At Year 1? In FAIR?

```{r}
summarize_results <- function(met_type, imp_symptoms, df_limma_lucid, df_limma_fair, df_ml_lucid, df_ml_fair, signif_only = TRUE) {
    df_limma_lucid <- df_limma_lucid %>%
        filter(metab_type==met_type, imputed_symptoms==imp_symptoms)
    df_limma_fair <- df_limma_fair %>%
        filter(metab_type==met_type)
    df_ml_lucid <- df_ml_lucid %>%
        filter(metab_type==met_type, imputed_symptoms==imp_symptoms) %>%
        left_join(unique(select(df_limma_lucid, metab_id, hmdb_id)), by = "metab_id")
    df_ml_fair <- df_ml_fair %>%
        filter(metab_type==met_type) %>%
        left_join(unique(select(df_limma_fair, metab_id, hmdb_id)), by = "metab_id")
    
    signif_combos_nonml <- lapply(list(df_limma_lucid, df_limma_fair), function(df) {
        df %>%
            filter(adj_pval < 0.1, !is.na(metabolite), !is.na(hmdb_id), hmdb_id!="NA") %>%
            select(symptom, hmdb_id)
    })
    signif_combos_nonml <- bind_rows(signif_combos_nonml)
    signif_combos_ml_lucid <- df_ml_lucid %>%
        mutate(
            is_imp_baseline = (importance_lasso_Baseline %in% c("High")) & (importance_rf_Baseline %in% c("High")),
            is_imp_year1 = (`importance_lasso_Year 1` %in% c("High")) & (`importance_rf_Year 1` %in% c("High"))
        ) %>%
        filter(is_imp_baseline | is_imp_year1) %>%
        select(symptom, hmdb_id)
    signif_combos_ml_fair <- df_ml_fair %>%
        filter(
            importance_lasso_FAIR %in% c("High"),
            importance_rf_FAIR %in% c("High")
        ) %>%
        select(symptom, hmdb_id)
    signif_combos <- bind_rows(signif_combos_nonml) %>% unique() %>% drop_na()
    if (!signif_only) {
        signif_combos <- df_limma_lucid %>%
            select(symptom, hmdb_id) %>%
            unique() %>%
            drop_na() %>%
            filter(hmdb_id != "NA")
    }
    
    ## Get all columns of relevant information
    limma_lucid_signif <- df_limma_lucid %>%
        semi_join(signif_combos, by = c("symptom", "hmdb_id")) %>%
        select(method, event, symptom, hmdb_id, metabolite, mz, rt, symptomlevel2, symptomlevel3, P.Value, adj_pval) %>%
        pivot_wider(names_from = event, values_from = c(symptomlevel2, symptomlevel3, P.Value, adj_pval))
    
    limma_fair_signif <- df_limma_fair %>%
        semi_join(signif_combos, by = c("symptom", "hmdb_id")) %>%
        select(method, symptom, hmdb_id, metabolite, mz, rt, former_nomenclature, symptomlevel2_FAIR = symptomlevel2, symptomlevel3_FAIR = symptomlevel3, P.Value_FAIR = P.Value, adj_pval_FAIR = adj_pval)
    
    ## Merge in FAIR
    res <- limma_lucid_signif %>%
        left_join(limma_fair_signif, by = c("method", "symptom", "hmdb_id"), suffix = c("_lucid", "_fair"))
    
    ## Clean up presentation
    res <- res %>%
        mutate(across(starts_with("symptomlevel"), function(x) {round(2^x, 2)})) %>%
        mutate(across(contains("adj_pval"), function(x) {ifelse(x < 0.1, paste("(S)", x), paste("", x))})) %>%
        arrange(symptom, adj_pval_Baseline) %>%
        left_join(df_ml_lucid, by = c("symptom", "metabolite_lucid" = "metab_id")) %>%
        left_join(df_ml_fair, by = c("symptom", "metabolite_fair" = "metab_id")) %>%
        select(method, symptom, hmdb_id, former_nomenclature, mz_lucid, rt_lucid, starts_with("metabolite_"), contains("Baseline"), contains("Year 1"), contains("FAIR"))

    res
}
```

```{r}
res_ruv <- summarize_results(
    met_type = "qrilc_ruv",
    imp_symptoms = TRUE,
    df_limma_lucid = df_limma_results_lucid,
    df_limma_fair = df_limma_results_fair,
    df_ml_lucid = df_ml_results_lucid,
    df_ml_fair = df_ml_results_fair,
    signif_only = TRUE
)

res_no_ruv <- summarize_results(
    met_type = "qrilc_no_ruv",
    imp_symptoms = TRUE,
    df_limma_lucid = df_limma_results_lucid,
    df_limma_fair = df_limma_results_fair,
    df_ml_lucid = df_ml_results_lucid,
    df_ml_fair = df_ml_results_fair,
    signif_only = TRUE
)

res_ruv_no_imp_symp <- summarize_results(
    met_type = "qrilc_ruv",
    imp_symptoms = FALSE,
    df_limma_lucid = df_limma_results_lucid,
    df_limma_fair = df_limma_results_fair,
    df_ml_lucid = df_ml_results_lucid,
    df_ml_fair = df_ml_results_fair,
    signif_only = TRUE
)

res_no_ruv_no_imp_symp <- summarize_results(
    met_type = "qrilc_no_ruv",
    imp_symptoms = FALSE,
    df_limma_lucid = df_limma_results_lucid,
    df_limma_fair = df_limma_results_fair,
    df_ml_lucid = df_ml_results_lucid,
    df_ml_fair = df_ml_results_fair,
    signif_only = TRUE
)

res_ruv <- unique(res_ruv) %>% mutate(analysis_version = "ruv_yes_imp_yes")
res_no_ruv <- unique(res_no_ruv) %>% mutate(analysis_version = "ruv_no_imp_yes")
res_ruv_no_imp_symp <- unique(res_ruv_no_imp_symp) %>% mutate(analysis_version = "ruv_yes_imp_no")
res_no_ruv_no_imp_symp <- unique(res_no_ruv_no_imp_symp) %>% mutate(analysis_version = "ruv_no_imp_no")
```

Store results from all sensitivity analyses.

```{r}
df_analysis_versions <- crossing(
    met_type = c("qrilc_no_ruv", "qrilc_ruv"),
    imp_symptoms = c(TRUE, FALSE)
) %>% 
    mutate(analysis_version = c("ruv_no_imp_no", "ruv_no_imp_yes", "ruv_yes_imp_no", "ruv_yes_imp_yes"))

df_results_all_versions <- lapply(seq_len(nrow(df_analysis_versions)), function(r) {
    met_type <- df_analysis_versions$met_type[r]
    imp_symptoms <- df_analysis_versions$imp_symptoms[r]
    res <- summarize_results(
        met_type = met_type,
        imp_symptoms = imp_symptoms,
        df_limma_lucid = df_limma_results_lucid,
        df_limma_fair = df_limma_results_fair,
        df_ml_lucid = df_ml_results_lucid,
        df_ml_fair = df_ml_results_fair,
        signif_only = FALSE
    )
    res <- res %>%
        unique() %>%
        mutate(analysis_version = df_analysis_versions$analysis_version[r])
}) %>% bind_rows() %>% unique()
```

```{r}
se_all_lucid <- merge_ses_across_injections(se_list)
se_all_fair <- merge_ses_across_injections(se_list_fair)

"HMDB0000317" %in% rowData(se_all_fair)$hmdb_id
```

Plots of symptom grade versus metabolite abundance.

```{r}
hmdb_ids_interest <- res_ruv %>%
    filter(!str_detect(metabolite_lucid, "acetaminophen")) %>%
    pull(hmdb_id) %>%
    unique()

rd <- rowData(se_all_lucid)
bool_metabs <- rd$hmdb_id %in% hmdb_ids_interest
cd <- colData(se_all_lucid)
bool_samples <- cd$event_name=="Year 1"

se_all_lucid_subs <- se_all_lucid[bool_metabs, bool_samples]
se_all_lucid_subs <- se_all_lucid_subs[!duplicated(rowData(se_all_lucid_subs)$hmdb_id),]
df_interest <- assay(se_all_lucid_subs, "log_abund") %>% t() %>% as.data.frame()
colnames(df_interest) <- rowData(se_all_lucid_subs)$hmdb_id

df_metab_names <- rowData(se_all_fair)[rowData(se_all_fair)$hmdb_id %in% hmdb_ids_interest, c("hmdb_id", "metabolite")] %>% as.data.frame() %>% unique()
new_colnames <- tibble(hmdb_id = colnames(df_interest)) %>%
    left_join(df_metab_names, by = "hmdb_id") %>%
    mutate(metabolite = ifelse(is.na(metabolite), "2-hydroxy-3-methylpentanoate", metabolite)) %>%
    pull(metabolite)

colnames(df_interest) <- new_colnames

symptom_names <- get_symptom_definitions()$short_descrip
df_interest <- bind_cols(as.data.frame(colData(se_all_lucid_subs)[,symptom_names]), df_interest)
df_interest <- pivot_longer(df_interest, cols = fatigue:bodily_pain, names_to = "symptom", values_to = "severity") %>%
    pivot_longer(cols = -c(symptom, severity), names_to = "metabolite", values_to = "abundance")

df_interest <- df_interest %>%
    mutate(
        symptom = recode(symptom, "fatigue" = "Fatigue", "pruritus" = "Pruritus", "anorexia" = "Anorexia", "nausea_vomiting" = "Nausea / vomiting", "daytime_sleepiness" = "Daytime sleepiness", "difficulty_concentrating" = "Difficulty concentrating", "bodily_pain" = "Bodily pain"),
        severity = str_remove(severity, "grade")
    ) %>%
    filter(!is.na(severity))

pdf("../results/figures/boxplots_abund_severity.pdf", width = 7*3.7, height = 7*4.4)
ggplot(df_interest, aes(x = abundance, y = severity, color = severity)) +
    geom_tufteboxplot() +
    facet_grid(metabolite ~ symptom) +
    theme_classic() +
    theme(text = element_text(size = 24)) +
    labs(x = "Metabolite abundance", y = "Symptom severity")
dev.off()
```

Organize linear regression results.

```{r}
df_linreg_results <- df_linreg_results_lucid %>%
    filter(metab_type != "orig", !is.na(hmdb_id)) %>%
    select(method, symptom, event, hmdb_id, metabolite, estimate, p.value, adj_pval, metab_type, imputed_symptoms) %>%
    pivot_wider(names_from = event, values_from = c(estimate, p.value, adj_pval)) %>% 
    left_join(select(df_linreg_results_fair, method, symptom, hmdb_id, metabolite, estimate, p.value, adj_pval, metab_type), by = c("method", "symptom", "hmdb_id", "metab_type"), suffix = c("_lucid", "_fair")) %>%
    mutate(analysis_version = case_when(
        metab_type=="qrilc_ruv" & imputed_symptoms ~ "ruv_yes_imp_yes",
        metab_type=="qrilc_ruv" & !imputed_symptoms ~ "ruv_yes_imp_no",
        metab_type=="qrilc_no_ruv" & imputed_symptoms ~ "ruv_no_imp_yes",
        metab_type=="qrilc_no_ruv" & !imputed_symptoms ~ "ruv_no_imp_no"
    )) %>%
    select(-metab_type, -imputed_symptoms, -metabolite_fair) %>% 
    rename(metabolite = metabolite_lucid, estimate_fair = estimate, p.value_FAIR = p.value, adj_pval_FAIR = adj_pval)

df_linreg_results_signif <- df_linreg_results %>%
    filter(adj_pval_Baseline < 0.1 | `adj_pval_Year 1` < 0.1 | adj_pval_FAIR < 0.1) %>%
    mutate(across(contains("adj_pval"), function(x) {ifelse(x < 0.1, paste("(S)", x), paste("", x))}))

df_linreg_results <- df_linreg_results %>%
    mutate(across(contains("adj_pval"), function(x) {ifelse(x < 0.1, paste("(S)", x), paste("", x))}))
```

## Unknowns

Exploring unknown metabolites in the linear regression results.

```{r}
plot_known_unknown <- function(df, yvar) {
    df %>% 
        mutate(known = !is.na(metabolite)) %>% 
        filter(adj_pval < 0.1) %>% 
        ggplot(aes(x = known, y = {{ yvar }})) +
            geom_boxplot()
}

plot_known_unknown(df = df_linreg_results_lucid, yvar = estimate)
plot_known_unknown(df = df_linreg_results_fair, yvar = estimate)

plot_known_unknown(df = df_linreg_results_lucid, yvar = adj_pval)
plot_known_unknown(df = df_linreg_results_fair, yvar = adj_pval)
```

Exploring unknown metabolites in the limma results.

```{r}
plot_known_unknown(df = df_limma_results_lucid, yvar = symptomlevel2)
plot_known_unknown(df = df_limma_results_fair, yvar = symptomlevel2)

plot_known_unknown(df = df_limma_results_lucid, yvar = symptomlevel3)
plot_known_unknown(df = df_limma_results_fair, yvar = symptomlevel3)

plot_known_unknown(df = df_limma_results_lucid, yvar = adj_pval)
plot_known_unknown(df = df_limma_results_fair, yvar = adj_pval)
```

Look at results tables for significant unknowns.

```{r}
df_linreg_results_lucid %>% 
    filter(is.na(metabolite), adj_pval < 0.1) %>% 
    arrange(desc(abs(estimate))) %>% 
    select(event, symptom, estimate, adj_pval, metab_id, metab_type, imputed_symptoms, perc_missing)

df_linreg_results_fair %>% 
    filter(is.na(metabolite), adj_pval < 0.1) %>% 
    arrange(desc(abs(estimate))) %>% 
    select(event, symptom, estimate, adj_pval, metab_id, metab_type, perc_missing)

df_limma_results_lucid %>% 
    filter(is.na(metabolite), adj_pval < 0.1) %>% 
    arrange(adj_pval, desc(pmax(abs(symptomlevel2), abs(symptomlevel3)))) %>% 
    select(event, symptom, symptomlevel2, symptomlevel3, adj_pval, metab_id, metab_type, imputed_symptoms, perc_missing)

df_limma_results_fair %>% 
    filter(is.na(metabolite), adj_pval < 0.1) %>% 
    arrange(adj_pval, desc(pmax(abs(symptomlevel2), abs(symptomlevel3)))) %>% 
    select(event, symptom, symptomlevel2, symptomlevel3, adj_pval, metab_id, metab_type, perc_missing)
```

Combine results for unknown metabolites.

```{r}
get_signif_unknowns <- function(df) {
    df %>% 
        filter(is.na(metabolite), adj_pval < 0.1) %>%
        select(mz, rt, symptom, method) %>%
        unique()
}

unknowns_signif_linreg_lucid <- get_signif_unknowns(df_linreg_results_lucid %>% filter(metab_type != "orig"))
unknowns_signif_linreg_fair <- get_signif_unknowns(df_linreg_results_fair %>% filter(metab_type != "orig"))
unknowns_signif_limma_lucid <- get_signif_unknowns(df_limma_results_lucid %>% filter(metab_type != "orig"))
unknowns_signif_limma_fair <- get_signif_unknowns(df_limma_results_fair %>% filter(metab_type != "orig"))

unknowns_signif <- bind_rows(unknowns_signif_linreg_lucid, unknowns_signif_linreg_fair, unknowns_signif_limma_lucid, unknowns_signif_limma_fair) %>% unique()

# For each metabolite in unknowns_signif, get the results for linreg and limma
df_linreg <- bind_rows(
    select(df_linreg_results_lucid %>% filter(is.na(metabolite)), method, mz, rt, event, symptom, estimate, p.value, adj_pval, metab_type, imputed_symptoms),
    select(df_linreg_results_fair %>% filter(is.na(metabolite)) %>% mutate(imputed_symptoms = FALSE), method, mz, rt, event, symptom, estimate, p.value, adj_pval, metab_type, imputed_symptoms)
)
# %>%
#     pivot_wider(names_from = event, values_from = c(estimate, p.value, adj_pval)) %>%
#     arrange(paste(mz, rt), symptom, metab_type, imputed_symptoms)

df_limma <- bind_rows(
    select(
        df_limma_results_lucid %>% filter(is.na(metabolite)),
        method, mz, rt, event, symptom, symptomlevel2, symptomlevel3, P.Value, adj_pval, metab_type, imputed_symptoms),
    select(
        df_limma_results_fair %>% filter(is.na(metabolite)) %>% mutate(imputed_symptoms = FALSE),
        method, mz, rt, event, symptom, symptomlevel2, symptomlevel3, P.Value, adj_pval, metab_type, imputed_symptoms)
)
# %>%
#     pivot_wider(names_from = event, values_from = c(symptomlevel2, symptomlevel3, P.Value, adj_pval)) %>%
#     arrange(paste(mz, rt), symptom, metab_type, imputed_symptoms)
```

For the LUCID runs (but not the FAIR runs), some unknown metabolites have different `compound_id`'s but the same m/z and RT.

```{r}
# Some metabolites have 2 p-values
df_linreg[["adj_pval_Baseline"]] %>% lengths() %>% table()
df_limma[["adj_pval_Baseline"]] %>% lengths() %>% table()

df_linreg[["adj_pval_Year 1"]] %>% lengths() %>% table()
df_limma[["adj_pval_Year 1"]] %>% lengths() %>% table()

df_linreg[["adj_pval_FAIR"]] %>% lengths() %>% table()
df_limma[["adj_pval_FAIR"]] %>% lengths() %>% table()


df_limma[lengths(df_limma[["adj_pval_Baseline"]])==2,]

rd_hilic_pos <- rowData(se_hilic_pos_main) %>% as.data.frame()

rd_hilic_pos %>% filter(mz==128.0699, rt==9.14)
```

```{r}
match_unknowns <- function(mz, rt, symptom, method) {
    this_mz <- mz
    this_rt <- rt
    this_symptom <- symptom
    this_method <- method
    filter_df <- function(df, count_only = TRUE) {
        df_clean <- df %>%
            filter(
                mz >= this_mz-0.05, mz <= this_mz+0.05,
                rt >= this_rt-0.25, rt <= this_rt+0.25,
                symptom==this_symptom,
                method==this_method,
                metab_type != "orig"
            )
        if (count_only) {
            df_clean %>%
                group_by(event) %>%
                summarize(any_signif = any(adj_pval < 0.1, na.rm = TRUE))
        } else {
            df_clean
        }
    }
    list(
        original = tibble(mz = mz, rt = rt, symptom = symptom, method = method),
        num_matches_limma = df_limma %>% filter_df(count_only = TRUE),
        num_matches_linreg = df_linreg %>% filter_df(count_only = TRUE),
        possible_matches_limma = df_limma %>% filter_df(count_only = FALSE) %>% select(-method, -symptom) %>% as.data.frame(),
        possible_matches_linreg = df_linreg %>% filter_df(count_only = FALSE) %>% select(-method, -symptom) %>% as.data.frame()
    )
}

unknowns_signif_matched <- pmap(unknowns_signif, match_unknowns)

bool_replicated_limma <- map_lgl(unknowns_signif_matched, function(l) {
    sum(l$num_matches_limma$any_signif) >= 2
})
bool_replicated_linreg <- map_lgl(unknowns_signif_matched, function(l) {
    sum(l$num_matches_linreg$any_signif) >= 2
})

table(bool_replicated_limma)
table(bool_replicated_linreg)

unknowns_signif_matched[bool_replicated_linreg]
```

Put replication information on the unknowns in a neater data frame.

```{r}
signif_events <- map(unknowns_signif_matched, function(l) {
    get_matches_str <- function(df) {
        if (!any(df$any_signif)) {
            "None"
        } else {
            df %>% filter(any_signif) %>% pull(event) %>% str_c(collapse = ", ")
        }
    }
    matches_limma <- get_matches_str(l$num_matches_limma)
    matches_linreg <- get_matches_str(l$num_matches_linreg)
    
    lucid <- l$possible_matches_limma %>% filter(event %in% c("Baseline", "Year 1")) %>% select(mz, rt) %>% unique()
    fair <- l$possible_matches_limma %>% filter(event == "FAIR") %>% select(mz, rt) %>% unique()
    
    mz_lucid <- lucid %>% pull(mz) %>% unique() %>% str_c(collapse = ", ")
    rt_lucid <- lucid %>% pull(rt) %>% unique() %>% str_c(collapse = ", ")
    mz_fair <- fair %>% pull(mz) %>% unique() %>% str_c(collapse = ", ")
    rt_fair <- fair %>% pull(rt) %>% unique() %>% str_c(collapse = ", ")
    
    len0_to_na <- function(x) {
        if (length(x)==0) {
            NA
        } else {
            x
        }
    }
    
    adj_pvals <- l$possible_matches_limma %>%
        filter(metab_type != "orig") %>% 
        group_by(event) %>% 
        summarize(min_adj_pval = min(adj_pval, na.rm = TRUE))
    adj_pval_baseline <- adj_pvals %>% filter(event=="Baseline") %>% pull(min_adj_pval) %>% len0_to_na()
    adj_pval_year1 <- adj_pvals %>% filter(event=="Year 1") %>% pull(min_adj_pval) %>% len0_to_na()
    adj_pval_fair <- adj_pvals %>% filter(event=="FAIR") %>% pull(min_adj_pval) %>% len0_to_na()
    
    res <- c(matches_limma, matches_linreg, mz_lucid, rt_lucid, mz_fair, rt_fair, adj_pval_baseline, adj_pval_year1, adj_pval_fair)
    names(res) <- c("limma_signif_groups", "linreg_signif_groups", "mz_lucid", "rt_lucid", "mz_fair", "rt_fair", "min_limma_adj_pval_baseline", "min_limma_adj_pval_year1", "min_limma_adj_pval_fair")
    res
}) %>% bind_rows()

unknowns_signif_rep_info <- bind_cols(unknowns_signif, signif_events) %>% 
    mutate(is_replicated = str_detect(limma_signif_groups, ",") | str_detect(linreg_signif_groups, ","))
```

Add adjusted p-value information.

```{r}
df_limma_wide <- df_limma %>%
    select(-c(symptomlevel2, symptomlevel3, P.Value)) %>% 
    mutate(event = str_remove(event, " ")) %>% 
    group_by(method, mz, rt, symptom, event, metab_type, imputed_symptoms) %>% 
    summarize(adj_pval = min(adj_pval)) %>% 
    ungroup() %>% 
    pivot_wider(id_cols = c(method, mz, rt, symptom), names_from = c(event, metab_type, imputed_symptoms), values_from = adj_pval)

temp <- unknowns_signif_rep_info %>%
    left_join(df_limma_wide %>% select(-contains("_orig_")), by = c("mz", "rt", "symptom", "method"))
```

Write results tables to file.

```{r}
# limma+ML results
# Only significant results
df_results_signif_all_versions <- bind_rows(res_ruv, res_no_ruv, res_ruv_no_imp_symp, res_no_ruv_no_imp_symp)
write_csv(df_results_signif_all_versions, file = "../results/tables/symptoms_results_signif_limma_ml_all_versions.csv")

# All results
write_csv(df_results_all_versions, file = "../results/tables/symptoms_results_all_limma_ml_all_versions.csv")

# OLS results
write_csv(df_linreg_results, file = "../results/tables/symptoms_results_all_linreg_all_versions.csv")
write_csv(df_linreg_results_signif, file = "../results/tables/symptoms_results_signif_linreg_all_versions.csv")

# Unknowns
write_csv(unknowns_signif_rep_info %>% select(-mz, -rt), "../results/tables/symptoms_results_unknowns.csv")
```


